---
title: "Task1"
author: "Dmytro Morozov"
date: "10/27/2025"
format:
  html:
    embed-resources: true
editor: source
---

## Question 1 Atomic Types

Create the following variables related to your chosen object:

4 numeric values (e.g., temperature, elevation, number of cars, highest building) 3 integer values (e.g., population, area, number of shopping malls) 2 logical values (e.g., has_water, is_inhabited) 3 character values (e.g., name, region, category)

My object is Kyiv city

```{r}
# numeric
temperature <- 7.0
meters_above_sea_level <- 131
auto <- 3020000
highest_building <- 141.0

# integer
population <- 3020228
area <- 839 # km²
shopping_malls <- 6

# logical
has_water <- TRUE
is_inhabited <- TRUE

# character
name <- "Kyiv"
region <- "Kyiv City"
category <- "City"
```

## Question 2: Vectors

Create three different vectors:

Numeric vector combining all your numeric and integer variables. (Use c().)

Sequence vector from any number between 10–50 to 100, choose step between 2 and 10. (Use seq() or :.)

Logical vector of length 10 with random TRUE/FALSE values using sample().

```{r}
# Numeric vector combining all numeric and integer variables
vec <- c(temperature, meters_above_sea_level, auto, highest_building, population, area, shopping_malls)

# Sequence vector from any number between 10–50 to 100, choose step between 2 and 10
vec_seq <- seq(from = 20, to = 100, by = 6)

# Logical vector of length 10 with random TRUE/FALSE values using sample()
vec_log <- sample(c(TRUE, FALSE), 10, replace = TRUE)
```

## Question 3: Matrix

Create a 3×3 matrix with random numbers between -50 and 50 (using sample() or runif()), and name its rows and columns according to your theme. Then calculate row and column averages using rowMeans() and colMeans().

```{r}
# Create a 3×3 matrix with random numbers between -50 and 50
m <- matrix(sample(-50:50, 9), nrow = 3, ncol = 3)

# Name its rows and columns according to your theme
rownames(m) <- c("District 1", "District 2", "District 3")
colnames(m) <- c("Metric A", "Metric B", "Metric C")

# Calculate row and column averages
row_means <- rowMeans(m)
col_means <- colMeans(m)
```

## Question 4: List

Create a list that combines all your previous objects: atomic variables, vectors, and matrix. Add one extra element — a short description of your universe as a string.

```{r}
# Create a list that combines all your previous objects
my_list <- list(
  numeric_values = c(temperature, meters_above_sea_level, auto, highest_building),
  integer_values = c(population, area, shopping_malls),
  logical_values = c(has_water, is_inhabited),
  character_values = c(name, region, category),
  numeric_vector = vec,
  sequence_vector = vec_seq,
  logical_vector = vec_log,
  matrix = m,
  description = "Kyiv is the capital and largest city of Ukraine, known for its rich history and cultural heritage."
)
```

## Question 5: Factor

Create a factor variable representing categories within your universe.

```{r}
# Create a factor variable representing categories within your universe
my_factor <- factor(c("Capital", "Large", "Historic", "Cultural", "Modern"), levels = c("Historic", "Cultural", "Modern", "Large", "Capital"))
```

## Question 6: Data Frame

Create a data frame related to your object that includes:

5–8 observations (rows)

At least 3 columns of different data types (e.g., name, numeric value, logical flag, category) Then:

Rename the columns using names() or colnames().

Display its structure with str() and its summary with summary().

Add a new column derived from the others (for example, a ratio or condition).

```{r}
# Create a data frame related to your object
data <- data.frame(
  name = c("Podil", "Pechersk", "Obolon", "Shevchenkivskyi", "Sviatoshynskyi"),
  population = c(150000, 200000, 250000, 180000, 220000),
  has_metro = c(TRUE, TRUE, TRUE, TRUE, FALSE),
  district_type = c("Historic", "Government", "Residential", "Commercial", "Industrial")
)

# Rename the columns
colnames(data) <- c("District Name", "Population", "Has Metro", "District Type")

# Display its structure
str(data)

# Display its summary
summary(data)

# Add a new column derived from the others
data$population_density <- data$Population / area
```

## Task 2: Querying and Subsetting Data

### 2.1 Querying Vectors

#### 2.1.1 Sequence Vector

```{r}
cat("\n2.1.1. Querying Sequence vector:\n")

# Display single element
print(vec_seq[3])

# Display range of elements
print(vec_seq[2:5])

# Display elements greater than 50
print(vec_seq[vec_seq > 50])
```

#### 2.1.2 Logical Vector

```{r}
cat("\n2.1.2. Querying Logical vector:\n")

# Show TRUE elements
print(vec_log[vec_log])

# Show indices of TRUE elements
print(which(vec_log))
```

---

### 2.2 Querying Your “Universe” List

```{r}
cat("\n2.2. Querying List:\n")

# Character values (from Task 1)
print(my_list$character_values)

# Show matrix
print(my_list$matrix)

# Show one numeric element
print(my_list$numeric_values[2])
```

---

### 2.3 Investigating Your Data Frame

```{r}
cat("\n2.3. Querying Data Frame:\n")

# Display first 3 rows
print(head(data, 3))

# Display one specific value
print(data[4, 1])

# Display logical column
print(data$`Has Metro`)

# Subset where population > 200000
df_subset <- data[data$Population > 200000, ]
print(df_subset)

# Subset for specific districts
df_subset_2 <- data[data$`District Name` %in% c("Pechersk", "Obolon"), ]
print(df_subset_2)
```

---

### 2.4 Handling “Corrupted” Data (NA)

```{r}
cat("\n2.4. Handling NA Data:\n")

# Create corrupted copy
df_corrupted <- data
df_corrupted[2, 3] <- NA
df_corrupted[5, 2] <- NA
df_corrupted[4, 5] <- NA

# Count NA values
print(sum(is.na(df_corrupted)))
print(colSums(is.na(df_corrupted)))

# Mean calculations
print(mean(df_corrupted$Population))
print(mean(df_corrupted$Population, na.rm = TRUE))

# Remove NAs
df_clean <- na.omit(df_corrupted)
print(df_clean)
```

---

### 2.5 Bonus Challenge: which()

```{r}
cat("\n2.5. Bonus: which.max() and which.min():\n")

# Find highest population district
highest_pop_index <- which.max(data$Population)
print(highest_pop_index)
print(data[highest_pop_index, ])
```