---
author: Dmytro Morozov
date: "11/25/2025"
editor: source
format:
  html:
    embed-resources: true
title: "Task 3. Urban Noise Analysis"
---
```{r setup, message=FALSE, warning=FALSE, include=FALSE}
library(dplyr) 
library(readr) 
urban_noise <- read_csv("urban_noise_levels.csv")
```

### Question 1: What are the 10 loudest recorded noise measurements?

```{r}
urban_noise %>%
  select(datetime, decibel_level, latitude, longitude) %>%
  arrange(desc(decibel_level)) %>%
  head(10)
```

### Question 2: How many measurements were recorded on weekends vs weekdays?

```{r}
urban_noise %>%
  group_by(is_weekend) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
```

### Question 3: Which hours of the day have the highest average noise level?

```{r}
urban_noise %>%
  group_by(hour) %>%
  summarise(avg_noise = mean(decibel_level, na.rm = TRUE)) %>%
  arrange(desc(avg_noise)) %>%
  head(5)
```

### Question 4: Create a new column showing noise-to-traffic ratio.

```{r}
urban_noise %>%
  mutate(noise_traffic_ratio = decibel_level / (vehicle_count + 1)) %>% 
  select(datetime, decibel_level, vehicle_count, noise_traffic_ratio) %>%
  arrange(desc(noise_traffic_ratio)) %>%
  head(10)
```

### Question 5: Which environmental factor (temperature) corresponds to the highest average noise levels?

```{r}
urban_noise %>%
  mutate(temp_group = cut(temperature_c,
                          breaks = c(-Inf, 5, 10, 15, 20, 25, Inf),
                          labels = c("very cold", "cold", "cool", "mild", "warm", "hot"))) %>%
  group_by(temp_group) %>%
  summarise(avg_noise = mean(decibel_level, na.rm = TRUE), count = n()) %>%
  arrange(desc(avg_noise))
```
